Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-03-30 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

\section{Model architecture and training strategy}
Self-Supervised Learning (SSL) is a technique to learn latent representations from large-scale data without relying on human annotations. Instead, it solves pretext tasks by creating different views of an object that are highly correlated. An SSL model is then trained to generalize objects' representations in a latent high-dimensional space. To achieve this, the model contrasts the representations of the same object with other objects (defined as negative samples) during training. By doing so, a contrastive SSL model is expected to produce more distinctive representations.

>>>>

\section{Model architecture and training strategy}
\subsection{Contrastive Self-Supervised Learning for Musical Representations (CSSLMR)}

Contrastive Self-Supervised Learning for Musical Representations (CSSLMR) is an approach that combines the principles of Self-Supervised Learning (SSL) and Contrastive Learning of Musical Representations (CLMR) to learn discriminative and valuable representations of music without relying on explicit labels. By leveraging the structure and content of music and contrasting different augmentations or transformations of the same musical piece against randomly (or not) selected pieces, CSSLMR models can effectively capture the underlying structure and known invariance in the data.

The CSSL methodology involves the following steps:

\textbf{Data Augmentation:} Create different augmentations or transformations of the same musical piece as positive pairs and use other randomly selected pieces as negative pairs.
\textbf{Contrastive Learning:} Train a neural network to minimize a contrastive loss function, which encourages the model to produce similar representations for positive pairs and dissimilar representations for negative pairs.
\textbf{Latent Representation:} Utilize the learned representations for various MIR downstream tasks, such as music genre classification or music similarity, both in the symbolic and time domains.
\subsubsection{Strengths}

CSSL inherits the strengths of both SSL and CLMR:

\begin{enumerate}
\item Reduces dependency on labeled data: By leveraging inductive bias and the structure of musical data, CSSL can learn from large amounts of unlabeled data.
\item Encourages feature extraction: CSSL promotes learning useful and discriminative features or representations
\end{enumerate}
